---
title: "Exercise 5"
output: html_notebook
---

```{r}
library(stringr)
library(dplyr)
```

Define a function, which given a numeric, integer or character vector, , identifying one of the CSV files, opens it and prints a of the data to the console if requested. Note that the name of each file begins with ‘part’ followed by the which is always represented by a three digit number. Do not use to write this function. The return value of the function is a data frame. Use the following scheme:

```{r}
getData <- function(directory, id, summarize = FALSE, verbose=FALSE) {
  # 'id' is a vector of length 1 indicating a dataset.
  # The user should be able to specify 'id' either as an integer, a
  # character, or a numeric value.
  
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'summarize' is a logical indicating if a summary of
  # the data should be printed to the console.
  fileName <- paste0("part", sprintf("%03d",as.numeric(id)), ".csv")
  filePath <- file.path(directory, fileName)
  if (verbose == TRUE) {
    print(paste("Reading:", gsub("//", "/",filePath)))
    }
  data <- read.csv2(file = filePath)
  if (summarize == TRUE) {
    print(summary(data))
  }
  data
}
```

```{r}
dat <- getData("data/Data Collection/", 
               id = 12, 
               summarize = TRUE)
```

# 2
2.Define a function which given a vector of id’s, will load the appropriate files and returns a data frame containing all loaded datasets. Create an integer variable in the return data which identifies the original dataset. Use the following scheme:

```{r}
appendData <- function(directory, id = 1:270) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  # create list to hold data frames with length equal to number of files to load
  listOfDataFrames <- vector("list", length(id))
  
  # loop in ids 
  for (fileId in id) {
    data <- getData(directory = directory, id = fileId, summarize = FALSE)
    data$id <- fileId
    listOfDataFrames[[fileId]] <- data
  }
  # Merge list of Data Frames into dfMerged
  print("Merging Data Frames")
  dfMerged <- do.call("rbind", listOfDataFrames)
}  
  
```
```{r}
dat <- appendData("data/Data Collection/", id = 1:3)
dim(dat)
```


## other way

```{r}
appendData2 <- function(directory, id = 1:270) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  # create list to hold data frames with length equal to number of files to load
  idChar <- formatC(id, width = 3, flag = 0)
  fileNames <- paste0(directory,"part",idChar,".csv")
  resList <- lapply(fileNames, read.csv2)
  do.call(rbind, resList)
}
```

```{r}
dat <- appendData2("data/Data Collection/", id = 1:5)
dim(dat)
```

```{r}
myRead.csv2 <- function(file, id) {
  res <- read.csv2(file)
  res[["id"]] <- id
  return(res)
}

appendData3 <- function(directory, id = 1:270) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  # create list to hold data frames with length equal to number of files to load
  idChar <- formatC(id, width = 3, flag = 0)
  fileNames <- paste0(directory,"part",idChar,".csv")
  resList <- mapply(FUN = myRead.csv2,
                    fileNames, id, 
                    SIMPLIFY = FALSE)
  resDF <- do.call(rbind, resList)
  rownames(resDF)
  return(resDF)
}
```

```{r}
dat <- appendData3("data/Data Collection/", id = 1:5)
dat
```


# 3 

3. Define a function which given a vector of id’s will load the corresponding datasets by using
the function appendData from the previous exercise. It will then calculate the correlation
between the variables carat and price and between volume and price, as well as the mean 
of carat, price and volume for each id! Volume can be approximated by x · y · z. Use the
following scheme:

```{r}
calculateStatistics <- function(directory, id = 1:270) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded.
  
  # Return a data frame with the correlations and means.
  
  appendData("data/Data Collection/", id = id) %>% 
    mutate(volume = x * y * z) %>% 
    group_by(id) %>% 
    summarise(corrCP = cor(carat, price),
              corrVP = cor(volume, price),
              meancarat = mean(carat),
              meanVolume = mean(volume),
              meanPrice = mean(price)
              )
}
```

```{r}
dat <- calculateStatistics("data/Data Collection/")
dat[1:3, ]
```

4. Rewrite the function appendData such that if the argument id = NULL the function will
try to open and append any file matching the pattern partXXX.csv in the directory given by
directory. The default value of id should be NULL for thi

```{r}
appendDataExtra <- function(directory, id = NULL) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  if (is.null(id)) {
    listOfFiles <- list.files("data/Data Collection")
    pattern <- "part[0-9][0-9][0-9].csv"
    
    # matches "partXXX.csv" and extract ids 
    id <- str_extract(str_subset(listOfFiles, pattern), "[0-9]+")
  } 
  
  # get number of files to be imported
  numberOfFiles <- length(id)
  
  # create list to hold data frames with length equal to number of files
  listOfDataFrames <- vector("list", numberOfFiles)
  
  # modify following loop for lapply
  fileNames <- paste0(directory, "part", as.character(id), ".csv")

  # loop in ids 
  for (fileId in id) {
    data <- getData(directory = directory, id = fileId, summarize = FALSE)
    data$id <- fileId
    
    # if only 1 file, return data, otherwise append data to a list
    if (numberOfFiles == 1) {
      return(data)
    }
    listOfDataFrames[[fileId]] <- data
  }
  # Merge list of Data Frames into dfMerged
  print("Merging Data Frames")
  dfMerged <- do.call("rbind", listOfDataFrames)
}
```


#### other way
```{r}
appendDataExtra2 <- function(directory, id = NULL) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  if (is.null(id)) {
    listOfFiles <- list.files("data/Data Collection")
    pattern <- "part[0-9][0-9][0-9].csv"
    
    # matches "partXXX.csv" and extract ids 
    id <- str_extract(str_subset(listOfFiles, pattern), "[0-9]+")
  } 
  
  # get number of files to be imported
  numberOfFiles <- length(id)
  
  # create list to hold data frames with length equal to number of files
  listOfDataFrames <- vector("list", numberOfFiles)
  
  # modify following loop for lapply
  fileNames <- paste0(directory, "part", as.character(id), ".csv")
  
  # define function for mapply
  myRead.csv2 <- function(file, id) {
    res <- read.csv2(file)
    res[["id"]] <- id
    return(res)
  }
  
  resList <- mapply(FUN = myRead.csv2,
                    fileNames, id, 
                    SIMPLIFY = FALSE)
  resDF <- do.call(rbind, resList)
  rownames(resDF) <- NULL
  return(resDF)
}
dat <- appendDataExtra2("data/Data Collection/")
dat
```



```{r}
dat <- appendDataExtra("data/Data Collection/")
range(dat$id)
```



```{r}
appendDataMoreExtra <- function(directory, id = 1:270, verbose=TRUE) {
  # 'directory' is a character vector of length 1 indicating
  # the location of the CSV files.
  
  # 'id' is an integer, numeric or character vector indicating 
  # the datasets to be loaded. The ids don't have to be a sequence.
  
  # Return a data frame containing all loaded data sets with
  # an additional integer variable "id". 
  
  # get number of files in directory
  listOfFiles <- list.files("data/Data Collection")
  listOfFiles <- str_subset(string = list.files("data/Data Collection"),
                            pattern = "\\.csv$")
  
  if (is.null(id)) {
    pattern <- "part[0-9][0-9][0-9].csv"
    # matches "partXXX.csv" and extract ids 
    id <- str_extract(str_subset(listOfFiles, pattern), "[0-9]+")
  } 
  # get number of files to be imported
  numberOfFiles <- length(id)
  
  if (verbose == TRUE) {
    print(paste("Reading", numberOfFiles, "of", length(listOfFiles) , "*.csv files from", directory))
  }
  
  # create list to hold data frames with length equal to number of files
  listOfDataFrames <- vector("list", numberOfFiles)
  
  # loop in ids 
  for (fileId in id) {
    data <- getData(directory = directory, id = fileId, summarize = FALSE)
    data$id <- fileId
    
    # if only 1 file, return data, otherwise append data to a list
    if (numberOfFiles == 1) {
      return(data)
    }
    listOfDataFrames[[fileId]] <- data
  }
  # Merge list of Data Frames into dfMerged
  if (verbose == TRUE) {
    print("Merging Data Frames")
  } 
  dfMerged <- do.call("rbind", listOfDataFrames)
}

dat <- appendDataMoreExtra("data/Data Collection/", id = 1:13)
```

## The √N law

```{r}
meanVarSdSe <- function(x) {
  ### your code
  result <- c(mean <- mean(x),
              var <- var(x),
              sd <- sd(x),
              se <- sd / sqrt(length(x)))
  return(result)
}
meanVarSdSe(1:100)
```

Now run the code. Explain the result. Extend the function definition of meanVarSdSe with the
argument ..., as is illustrated as follows:
```{r}
meanVarSdSe(c(NA, 1:00))
```

```{r}
meanVarSdSe <- function(x, ...) {
  ### your code
  result <- c(N = as.integer(sum(!is.na(x))),
              mean = mean(x, ...),
              var = var(x, ...),
              sd = sd(x, ...),
              se = sd(x, ...) / sqrt(sum(!is.na(x))))
  return(result)
}
meanVarSdSe(c(NA, 1:100), na.rm=TRUE)
```


### Convergence and Control Structures

The N Law: the precision of the sample average improves with the square root of the sample size
N . Simulate 10 6 Poisson distributed random numbers with expectation value λ = 100 via

```{r}
set.seed(1) # why?
rPois <- rpois(n = 1e6, lambda = 100)
```


```{r}
# init 
n <- 2
se <- 1

result <- matrix(nrow = 0, ncol = 5)
while (se > 0.05) {

  # this causes errror when n > length(rPois)
  x <- rPois[1:n]
  
  # result 
  row <- meanVarSdSe(x)
  result <- rbind(result, row)

  # update
  n <- n * 2
  se <- row["se"]
  if (n > 3e+7) {
    break
  }
}

result
```

```{r}
x <- result[, 3]

plot(x, type = "l")
```





## useful links
https://stackoverflow.com/questions/2969315/rhow-to-get-grep-to-return-the-match-rather-than-the-whole-string
https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html