---
title: "Block 04 Apply Family"
author: "Marcelo Avila"
date: "January 8, 2018"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Apply Family

Writing for, while loops is useful when programming but not particularly easy when
working interactively on the command line. There are some functions which implement
looping to make life easier.

lapply
~ Loop over a list and evaluate a function on each element
sapply
~ Same as lapply but tries to simplify the result
mapply
~ Multivariate version of lapply
An auxiliary function split is also useful, particularly in conjunction with lapply.

### lapply()

```{r}
x <- list("a" = 1:5, 
          "b" = rnorm(10))

lapply(x, mean)
sap <-sapply(x, mean)
```

```{r}
x <- list(a = c(1:4, NA),
          b = rnorm(10),
          c = rnorm(20, 1),
          d = rnorm(100, 5)
)

lapply(x, mean, na.rm = TRUE)
```


```{r}
x 
```

```{r}
x <- 1:4
lapply(x, runif, min=0, max=100)
# same as runif( 1.. 1:2.. 1:3.. 1:4.., min=0, max=100)

# not like the following
for (i in x) {
  print(i)
}
```


```{r}
x <- 1:4
lapply(x, runif, min=0, n = 5)
```

```{r}
erg <- lapply(x, runif, min = 0, n = 1000)
lapply(erg, max)
```

```{r}
x <- list(a = matrix(data = 1:4,
                     nrow = 2,
                     byrow = F),
          b = matrix(data = 1:6,
                     nrow = 3,
                     byrow = TRUE))
x
```
```{r}
mean(x)
```


```{r}
lapply(x, function(mat) mat[ ,1])
```

### sapply()


```{r}
lapply(x, function(mat) {
  erg <-summary(mat[ ,2])
  return(erg)
})
```

sapply will try to simplify the result of lapply if possible.

- If the result is a list where every element is length 1, then a vector is returned.
- If the result is a list where every element is a vector of the same length (> 1),
  a matrix is returned.
- I If it can’t figure things out, a list is returned.
- I Check the manual for the `sapply` argument `simplify`.

```{r}
sapply(x, function(mat) {
  erg <-summary(mat[ ,2])
  return(erg)
})
```

#### split()

split takes a vector or other objects and splits it into groups determined by a factor or
list of factors.

```{r}
str(split)
```

x 
~ is usually a data frame
f
~ is a factor (or coerced to one) or a list of factors
drop
~ indicates if empty factor levels should be dropped
...
~ further potential arguments passed to methods.


```{r}
spl <- split(x <- airquality[, c("Ozone",
                                "Solar.R",
                                "Wind")],
             f <- airquality$Month)
```

```{r}
sapply(spl, colMeans)
```

```{r}
sapply(spl, colMeans, na.rm = TRUE)
```

```{r}
dataList <- lapply(split(x, f), quantile, na.rm=TRUE) #Split-Apply
do.call(cbind, dataList) #Combine the result
```
```{r}
sapply(split(x, f), quantile, na.rm=TRUE) #Split-Apply -> sapply ?
```



```{r}
testdf <- data.frame(x = rnorm(10),
                     fakt = factor(rep(1:2, 5)))

testdf
```

```{r}
```











### mapply()

```{r}
names <- c("Adam Riese", "Albert Einstein", "Oliver Kahn")

namesList <- strsplit(names, split = " ")
namesList <- lapply(namesList,
                    function(x) data.frame(
                      firstName = x[[1]],
                      lastName = x[[2]]
                      )
                    )

do.call(rbind, namesList)
```

```{r}
lHard <- list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
lSmart <- mapply(rep, 1:4, 4:1)
identical(lHard, lSmart)
```

```{r}
noise <- function(n, mean, sd) {
  rnorm(n, mean, sd)
}

noise(5, 1, 2)
noise(1:4, 1:4*10, 2)
```

```{r}
set.seed(1)
mapply(noise, 1:4, 1:4 * 10, 2)

set.seed(1)
list(noise(1, 10, 2), noise(2, 20, 2), noise(3, 30, 2), noise(4, 40, 2))
```
```{r}
noise <- function(n, mean, sd) {
  mapply(rnorm, n, mean, sd)
}

set.seed(1)
noise(1:4, 1:4 * 10, 2:3)
```

### apply 

```{r}
mat <- matrix(data = rnorm(1000),
              nrow = 200)

dim(mat)
rowSums(mat)

test <- apply(mat, MARGIN = 1, FUN = sum) 

length(test)

apply(mat, MARGIN = 2, FUN = sum)
apply(mat, MARGIN = 2, FUN = mean)

ar <- array(data = rnorm(4000), dim = c(200, 5, 4))

apply(ar, MARGIN = c(1, 2), FUN = mean)
dim(apply(ar, MARGIN = c(1, 2), FUN = mean))

apply(ar, MARGIN = 3, FUN = mean)



```

### tapply() Table Apply 

and by()





