---
title: "block4"
author: "Marcelo Avila"
date: "January 8, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Control Structures 

Control structures in R allow you to control the flow of execution of the program,
depending on runtime conditions. Common structures are
  if, else: testing a condition
  for: execute a loop a fixed number of times
  while: execute a loop while a condition is true
  repeat: execute an infinite loop
  break: break the execution of a loop
  next: skip an interation of a loop

```{r}
x <- c("a", "b", "c", "d")

for (i in seq_along(x)) {
  print(x[i])
}
```
```{r}
for(letter in x) {
  print(letter)
}
```
### Nesting
```{r}
x <- matrix(data = 1:6, ncol = 2, byrow = TRUE)

for(i in seq_len(nrow(x))) {
  for(j in seq_len(ncol(x))) {
    print(x[i, j])
  }
}
```

### AVOID
```{r}
x <- 1
for(i in 2:10) {
  x<- c(x, i)
}
```
### better
```{r}
x <- vector(length = 10)
for(i in 1:10) {
  x[i] <- i
}
x
```

## while
```{r}
number <- 0
while(number < 10) {
  print(number)
  number <- number + 1
}
number <- 0
while(number < 10) {
  number <- number + 1
  print(number)

}
```

```{r}
z <- 5
while(z >= 3 && z <=7) {
  print(z)
  coin <- rbinom(n = 1, size = 1, prob = 0.5)
  if(coin == 1) {
    z <- z + 1 
  } else {
    z <- z - 1
  }
}
```

```{r eval=FALSE}

x0 <- 1
tol <- 1e-8

repeat {
  x1 <- computeEstimate(x0)
  if(abs(x1 - x0) < tol) {
    break
  } else {
    x0 <- x1
  }
}
```

```{r}
for(i in 1:5) {
  if(i <= 2) {
    # hier soll wirklich nichts passieren
  } else {
    print(i)
  }
}

for(i in 1:5) {
  if(i <= 2) {
    next
  } else {
    print(i)
  }
}
```

```{r}

```

