---
title: "S3 Prog mit R - Class"
author: "Marcelo Avila"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## motivation 

Let’s assume we want to define a mean for the character type in R.
As mean we simply define the mean length of each element in a
vector of type character. The length of an element can be defined as
the number of characters.

```{r}
characterVector <- c("some", "more", "text", "and", "different", "nchar")
meanCharacter <- function(x) mean(nchar(x))
meanCharacter(characterVector)
```

```{r}
mean <- function(x) {
  mean(nchar(x))
} 
# name conflict

# base::mean geht kapput 
mean <- function(x) {
  base::mean(nchar(x))
}
mean(characterVector)
mean(1:10)

# create conditions
mean <- function(x) {
  if (is.character(x)) base::mean(nchar(x))
  else base::mean(x)
}

mean(characterVector)
mean(1:10)
```

There is one obstacle with the definition of meanCharacter, the name.
With every function we define we have to memorize one more function
name and also we have to find new useful names (which is not easy!).

How is a potential user (you in 2 weeks) ever going to understand and
know about all the functions you defined? Not at all.

If the name is a problem let’s define a new mean function.
One thing to keep in mind, however, is that we should preserve the
behaviour of the original mean function.

Every time we want to define a mean function for a new data type we
have to add more if clauses to the definition. In the long run this
strategy is going to be a mess.

Also we can not extend functionality, we have to change it. Extending
is good, changing is bad.


# The S3 class method: Basic Idea

The basic idea of the S3 class system is that it should be possible to
extend the functionality of a generic vocabulary without adding new
words (function names).

This is a convenience. You can throw any statistical model and data
type into the summary function. And somehow (allmost) for all data
types (linear models, data frames, etc.) summary knows what to do.

Even more, summary knows about data types from different packages,
although the original author had no way of anticipating these types.
To understand this we need to answer the three following questions:

What is a S3 class?

What is a generic function?

What is a method?


## What is a S3 class?

S3 classes serve a simple purpose, to give different or new data types
a name.

This is done by adding an attribute to the data with the name of the
class.


```{r}
dat <- list(c(1, 2), c(2, 3), c(2, 4))
class(dat) <- "rational"
str(dat)
```

Typically you do not assign the class interactively but you return data
with a class attribute from a constructor function (e.g. numeric, list,
lm).



```{r}
rational <- function(num, denom) {
  rat <- mapply(c, num, denom, SIMPLIFY = FALSE)
  class(rat) <- "rational"
  rat
}
```

```{r}
str(rational(c(1, 2, 2), c(2, 3, 4)))
```


rational is a constructor function for instances of class rational.
It is named constructor because it knows how to construct an object of
class rational.

There is no formal definition of the class, you simply say a list is of
class X.

In most scenarios the list type is used as the basic data structure to
compose new data (e.g. data.frame and lm).

So what is the great benefit of defining new classes? An example:
Printing an object of class rational to the console results in verbose
information. Now we can fix this:

```{r}
print.rational <- function(x, ...) cat(sapply(x, paste, collapse = "/"))
rational(c(1, 2, 2), c(2, 3, 4))
```


## Generic functions and methods

A generic functions is a function which only purpose is to find the
appropriate method given the class of its first argument.

The overall purpose is, that you have a function name, say mean, and
this function somehow figures out how the mean is defined for a given
data type.

The function print and mean are generic functions. print will find the
correct print method to print things to the console. And mean searches
for the correct mean method for a given data type.
Methods are defined by a naming convention: <generic>.<class>

```{r}
print.rational <- function(x, ...) {
  cat(sapply(x, paste, collapse = "/"))
}

mean.rational <- function(x, ...) {
  mean(sapply(x, function(el) el[1] / el[2]))
}
```

```{r}
rat <- rational(num, denom)

## was macht mean(sapply... )


out <- list()
for (i in 1:length(rat)) {
  out[[i]] <- rat[[i]][1] / rat[[i]][2]
}
out
```


```{r}
num <- c(1,2,2)
denom <- c(2,3,4)
```

```{r}
rational(num, denom)
mean(rational(num, denom))
```

We say print.rational is the print method for objects of class
rational. There is a print method for most of the data types in R which
define how output is printed to the console.





```{r}
# back to character example
mean(characterVector)

mean.character <- function(x, ...) {
  base::mean(nchar(x))
}

mean(characterVector)
```

the print function finds 

# ddtt 

## S3

- class
- method
- genereic
- object

class

- "bauplan"
     - constructor: erzeugt eine Instanz der klasse
     
```{r}
myClass <- function(x, y) {
  out <- list(x, y)
  class(out) <- "myClass"
  out
}


myClass(1, 2)
```

## Generic

```{r}
print <- function(x, ...) {
  UseMethod("print")
}
```


## method for "myClass"
```{r}
print.myClass <- function(x , ...) ...
```

```{r}
print.default <- function(x, ...) {
  
}
```

```{r}
# ?.S3methods
```

