---
title: "Block 03 - Data Handling"
author: "Marcelo"
date: "20 November 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
#rm(list = ls())
library(dplyr)
library(nycflights13)
```

dplyr aims to provide a function for each basic verb of data manipulating:

* filter() (and slice())
* arrange()
* select() (and rename())
* distinct()
* mutate() (and transmute())
* summarise()
* sample_n() and sample_frac()

```{r environment}
class(flights)
```

```{r}
dfFlights <- as.data.frame(flights)
```

```{r}
flights[flights$month == 7 &
          flights$day == 1,  ]
```

```{r}
system.time(filter(flights, month == 7, day == 1))
```

```{r}
system.time(filter(dfFlights, dep_delay < 0 & arr_delay > 0))
```


```{r}
#arrange(flights, year, month, day)
arrange(flights, desc(arr_delay))
```

all the following outputs are identical
```{r}
select(flights, year, month, day) 
select(flights, year:day)
select(flights, 1:3)
```


## Excercises: Subset, Filter and Arrange data frames

### 1 and 2
```{r}
dataPath <- "./Datasets/Schweiz/"
filesToLoad <- list.files(path = dataPath, full.names = TRUE)
filesToLoad
```

### 3
```{r}
dat1 <- read.csv2(file = filesToLoad[1])
str(dat1, vec.len = 1)
```


```{r}
dat2 <- read.csv(file = filesToLoad[2])
```



```{r}
dat1 <- read.table(file= filesToLoad[1],
                   header = TRUE,
                   sep = ";",
                   dec = ",",
                   stringsAsFactors = T)
head(dat1)
 
```

```{r}
dat2 <- read.table(file= filesToLoad[2],
                   header = TRUE,
                   sep = ",",
                   dec = ".",
                   stringsAsFactors = T)
head(dat2)
```



```{r}
dat3 <- read.table(file= filesToLoad[3],
                   header = TRUE,
                   sep = "_",
                   dec = "-",
                   stringsAsFactors = T)
head(dat3)
```

```{r}
dat4 <- read.table(file= filesToLoad[4],
                   header = TRUE,
                   sep = "\t",
                   dec = ",",
                   stringsAsFactors = T)
head(dat4)
```

```{r}
all(identical(dat1, dat2),
    identical(dat3, dat4),
    identical(dat1, dat3))
```


```{r}
datSub1 <- dat1[-(1:3), 1:6]
datSub2 <- dat1[-(4:10), c(1, 7)]
```


### 9.
Use the * join functions from the dplyr package to reunite the two subsets to one data frame
named datJoin. Read the help file for inner join. Which column should be used for joining?
Your result should look like:

```{r}
datJoin <- full_join(x = datSub1,
                      y = datSub2,
                      by = "Province")
str(datJoin, vec.len = 1)
identical(dim(dat1), dim(datJoin))
identical(dat1, datJoin)
```



## Subset, Filter and Arrange data frames

###


```{r}
data(swiss)
?swiss
```

### Create a new column swiss$Province in which you store the name of the provinces corresponding to the observations. Delete the rownames afterwards.

```{r}
swiss <- swiss %>%
  mutate(Province = rownames(swiss)) %>%
  select(7, 1:6)
swiss
``` 

### Identify the five provinces with the lowest percentage of males involved in agriculture as occupation. 


```{r}
swiss %>%
  arrange(Agriculture) %>%
  select(Province, Agriculture) %>%
  slice(1:5)
```



## Find the five provinces with the highest percentage of catholic population.
```{r}
swiss %>%
  arrange(desc(Catholic)) %>%
  select(Province, Catholic) %>%
  slice(1:5)
```

## Data Manipulation

```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(dplyr)
```



```{r}
urlRedWine <-
"http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv"
urlWhiteWine <-
"http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv"
readLines(urlRedWine, n = 5)
```

```{r}
readLines(urlWhiteWine, n = 5)
```

```{r}
redWine <- read.table(file = urlRedWine,
                      header = TRUE,
                      sep = ";",
                      dec = ".")

whiteWine <- read.table(file = urlWhiteWine,
                      header = TRUE,
                      sep = ";",
                      dec = ".")
```

```{r}
redWine <- mutate(.data = redWine, color = "red")
whiteWine <- mutate(.data = whiteWine, color = "white")
```

```{r}
wine <- bind_rows(redWine, whiteWine)
str(wine, vec.len = 2)
rm(redWine, whiteWine)
```

```{r}
# apply condition
wine <- wine %>%
  mutate(alcoholClass = ifelse(test = alcohol < 10,
                               yes = "low",
                               no = ifelse(test = alcohol < 12,
                                           yes = "medium",
                                           no = "high")))

# apply ordered factor
wine$alcoholClass <- factor(x = wine$alcoholClass,
                             levels = c("low", "medium", "high"),
                             ordered = TRUE)

#display aggregated table
table(wine$alcoholClass, wine$color)
```

```{r}
prop.table(table(wine$alcoholClass, wine$color), margin = 2)
```

### 6. Figure out how the same results can be obtained using dplyr functions:

```{r}
wine %>%
  select(color, alcoholClass) %>%
  count(color, alcoholClass)
```


```{r}
# not quite right
wine %>%
  select(color, alcoholClass) %>%
  count(color, alcoholClass) %>%
  group_by(color) %>%  mutate(ratio = n / sum(n))
```



## 7. Calculate the mean values for alcohol, density and pH for each combination of quality and color using the aggregate function.

```{r}
wineAgg <- wine %>%
  select(quality, color, alcohol, density, pH)
  
```


```{r}
aggregate(x = wineAgg[,-(1:2)],
          by = list(wineAgg$quality, wineAgg$color),
          FUN = mean)
```


















