---
title: "Programmieren mit R: Block 01"
output:
  html_document:
    toc: true
    toc_depth: 3
---

# Introduction

## Course Material

1. Roger Peng, Computing for data analysis, John Hopkins Uni
    a. available at github.com/DataScienceSpecialization
2. Chamber, Programming with R
3. ...


## Getting Help

1. stackoverflow
2. r-bloggers
3. rseek 
4. google


## What you have to do

i. exercise!
i. 5 informal + 3 formal (in groups)
i. find a homework group. NO more than 2-3 students. Send the names 
    and "Matr.Nr." to: Soeren.Pannier@fu-berlin.de before the deadline (??)


### grade:

i. 15% formal exercises (group)
i. 85% term paper (10 - 20 pages) (individually)


## what is a package?

i. set of functions and data sets which is not included in the "base" R system
ii. additional functionality

# Fundamentals

## Data types

R has six basic or "atomic" *data types*:

i. character
i. numeric
i. integer
i. complex
i. logical
i. raw

The most basic object is a vector

i. a vector can **only** contain objects of the same type  
    **Exception**: a list can contain objects of different classes

Data Frame: 
a list with equally long lists(columns). In each column only one atomic data type allowed 


```{r}
x <- 1
msg <- "hallo" 
y <- 3 * 4

# this is a comment.
z <- x:y
length(x); length(y); length(z); length(msg)
class(x); class(y); class(z); class(msg)

# combine function

list1 <- c(.5, .6, msg)     # character
list2 <- c(1,2,3)           # numeric
list3 <- c(TRUE, FALSE)     # logical
list4 <- c(T,F)
list5 <- c(1 + 2i, 2 + 4i)  # complex

identical(list3, list4)     # TRUE: lists 3 and 4 are identical
```

```{r}
x <- vector("numeric", length = 10)
x 
x <- vector(length = 10)
class(x)


y <- c(TRUE, 2)
class(y)

a <- 0:10
class(a)

as.numeric(a)
as.character(a)
as.complex(a)

f <- as.numeric(c("a", "b", "c"))
f
```

-- 02. Veranstaltung 23.10.2017 --

**Factors** are used to represent categorical data. Factors can be unordered or ordered. Internally factor are stored as "*labeled*" integer vectors.
> treated specially by modelling functions like lm() and glm()
> better self-describing than using intergesrs (e.g. Male = 0, Female = 1)

```{r}
relFac <- factor(c("christ", "christ", "jud", "isl", "isl"))
relFac
```



Defining levels explicitly. Non-matching values becomes <`NA`>
```{r}
x = factor(c("no", "yes", "no", "yes", "nein"),
           levels = c("yes", "no"))

table(x)

unclass(x)

relevel(x, ref = "no")
```

```{r}
exam <- c(TRUE, FALSE, TRUE, TRUE, FALSE)
factor(exam)

exam <- factor(exam,
               levels = c(TRUE, FALSE),
               labels = c("seccess", "failure"))
exam
```


### Data Frames

Data frames are used to store tabular data:

- They are represented as a special type of list where every element of the list has 
to have the same length.
- Each element of the list can be thought of as a column and the length of each
element of the list is the number of rows
- data frames can store different classes of objects in each column

x <- data.frame(eggs = 1:4,
ham = c(TRUE, TRUE, FALSE, FALSE))
x
```{r}
x <- data.frame(eggs = 1:4,
ham = c(TRUE, TRUE, FALSE, FALSE))
x

nrow(x)
ncol(x)
dim(x)
```


```{r}
a <- 1:12
b <- c("a", "b", "b")
d <- c(TRUE, TRUE, FALSE, FALSE)

x <- data.frame(fritz = a, "hanz" = b, d)
str(x)
```
```{r}
x <- list(a = 1, b = list(1, 2), c = data.frame(x = 1:2))
str(x)

```

### The NULL object

There is a special object called NULL. It

- represents the null object in R
- is an object with defined neutral ("null") behavior.
- has no type and no modifiable properties
- should not be confused with a vector or list of zero length
- is a reserved word
- is often returned by expressions and functions whose value is undefined
To test for NULL use `is.null()` (and not `a == NULL`)



## Subsetting

There are a number of operators that can be used to extract subsets of R objects.

- `[ ]` always returns an object of the same class as the original (there is one
exception); can be used to select more than one element
- `[[ ]]` is used to extract elements of a list or a data frame; it can only be used to
extract a single element and the class of the returned object will not necessarily be
a list or data frame
- `$` is used to extract elements of a list or data frame by name; semantics are similar
to that of `[[ ]]`.

```{r}
x <- c("a", "b", "c", "c", "d", "a")
x[1]
x[2]
```


Subsetting data frames with names: (slide 42)
```{r}
x <- data.frame(eggs = 1:4,
ham = c(TRUE, TRUE, FALSE, FALSE))
x$eggs

x[["eggs"]]

x[, "eggs"]

x[, "eggs"] ## integer
x[, "eggs", drop = F] ## data.frame

```

Subsetting data frames with positions: (slide 43)
```{r}
x <- data.frame(eggs = 1:4,
ham = c(TRUE, TRUE, FALSE, FALSE))
x[[1]] #class(x[[1]]) is integer

x[1] #class(x[1]) is data.frame

x[c(1,3), 2]
```

Subsetting data frames with logicals: (slide 44)
```{r}
x <- data.frame(eggs = 1:4,
ham = c(TRUE, TRUE, FALSE, FALSE))
x[x$eggs > 2, ]

x[x$ham, ] # because x$ham already T, T, F, F

x[c(TRUE, FALSE)]
```

To delete elements in a list you can simply assign the value NULL to them.
```{r}
x[1] <- NULL
x
```


For a subset of rows select the rows to keep or use data.frame[-<rows>, ]:

```{r}
x <- x[-(1:2), ]
x
```

### Missing Values

Missing values are denoted by `NA` or `NA` for undefined mathematical operations.

- `is.na()` is used to test objects if they are `NA`
- `is.nan()` is used to test for `NaN`
- `NA` values have a class also, so there are integer `NA`, character `NA`, etc.
- A `NaN` value is also `NA` but the converse is not true


```{r}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x)
bad
```


```{r}
x[!bad] # x filter >> not bad
```


```{r}
airquality[1:6,]
```

Removing NA values
```{r}
na.omit(airquality[1:6,]) # only four rows
```

```{r}
na.omit(airquality)[1:6,] # six rows. 
```

```{r}
identical(airquality[1:2], airquality[,1:2])
```


### Subsetting Lists

- Semantics for subsetting a list is equivalent to data frames because they are
closely related.
- One exception: A list is not meant to represent tabular data, hence subsetting
rows/columns is not meaningfull: list[<rows>, <cols>] won't work!
- Lists can be recursive (elements of a list can be lists). The subsetting for nested
elements works just like for a normal list:


```{r}
x <- list(a = list(10, 12, 14),
b = c(3.14, 2.81))
x[[1]][[3]]
```


```{r}

```











## Operators and Vectorized Operations







