---
title: "Programmieren mit R - Excercises"
output: html_notebook
---
# Ex1-CompStat:R
https://lms.fu-berlin.de/bbcswebdav/pid-3382855-dt-content-rid-5536805_2/courses/WIWISS_V_104063b_17W/Ex1.html


## Misc

1. Define the following vectors in R:
        a = sequence from 0 to 8
        b = (1, 1, 1, 2, 2, 2, 3, 3, 3)
        c = (1, 2, 2, 3, 3, 3, 4, 4, 4)
        d = (1, 2, 2, 3, 3, 3, 4, 4, 4)
        
```{r}
a = 0:8
b = c(1, 2, 2, 3, 3, 3, 4, 4, 4)
c = c(1, 2, 2, 3, 3, 3, 4, 4, 4)
d = c(1, 2, 2, 3, 3, 3, 4, 4, 4)
```

2. what are the dimensions and lengths of a, b, c and d?
```{r}
length(a)
## vector 1 x 9
```

3. Create a named vector namedNumbers containing the elements (1, 2, 3, 4) with element names (alpha, beta, gamma and delta)!

```{r}
namedNumbers = 1:4
names(namedNumbers) = c("alpha", "beta", "gamma", "delta")
namedNumbers
```


4. a, b, c and d are defined as before. Check the R Language Definiton 3.1.4, if you are not familiar with the applied operators (e.g. &, |, ==, etc.). Think first, check your results with R. What is the result of:



```{r}
1 == 2              # false
1 != 2              # true
1 == 2 & 2 == 2     # false
1 == 2 | 2 == 2     # true
```
```{r} 
"1" == 1        # false (! doch TRUE)
TRUE == 1       # true
FALSE == 0      # true
TRUE != FALSE   # true
```

```{r}      
TRUE & TRUE             # true 
FALSE | TRUE            # true
(TRUE | FALSE) & TRUE   # true 
(TRUE | FALSE) & FALSE  # true (DOCH false!)
```

```{r}
!(TRUE | FALSE)             # false
is.logical(c(T, F, T, F))   # true
is.logical(a)               # false
is.numeric(b)               # true
```

```{r}
is.numeric(A)   # ERROR 
```

```{r}
b == 2 # logical vector with TRUE where b = 2 and false otherwise
```

```{r}
any(b == 2)     # True
all(b == 2)     # false 
```

5. Extract only the first element of namedNumbers, use the indexing-function [. First use the position directly, second use the name and third use a logical vector!
```{r}
# position
namedNumbers[1]
# name
namedNumbers["alpha"]
#logical
namedNumbers[c(TRUE, FALSE, FALSE, FALSE)]
```

6. Extract only the elements of a, b, c and d which are equal to 1, 2, 3 and 4 respectively!
```{r}
longlist <- c(a[a > 0 & a < 5],
              b[b > 0 & b < 5],
              c[c > 0 & c < 5],
              d[d > 0 & d < 5])
longlist
```

7. What is the result of (think first):

```{r}
c("c", "1") > 1 # TRUE FALSE
```

```{r}
c("c", "1") > "b" # TRUE FALSE
```

```{r}
c("c", "1") > b # ERROR (longer object length is not a multiple of shorter object length)
```

```{r}
c("c", "1") > b[-1] # T, F, T, F, T, F, T, F, 
```
```{r}
c("c", "1") + c(TRUE, FALSE)
# ERROR (non-numeric argument to binary operator)
```

8. Calculate the mean, standard deviation, min, max and median for a, b, c and d!

```{r}
print("a")
mean(a); sd(a); max(a); median(a)
print("b")
mean(b); sd(b); max(b); median(b)
print("c")
mean(c); sd(c); max(c); median(c)
print("d")
mean(d); sd(d); max(d); median(d)

```

9. Which of the following names can be used for naming objects in R: A, A1, 1A, A=1, A&1,
.A, A.1, A_1, A 1

```{r}
# all of them (using quotation mark)
```

10. What are the general rules for naming objects in R? See the R-Language Definition 
for details.

    File Names
          
            File names should end in .R and, of course, be
            meaningful.
            
    GOOD: predict_ad_revenue.R
            
    BAD: foo.R
          

          Identifiers
          
            Don't use underscores ( _ ) or hyphens
            ( - ) in identifiers.
            Identifiers should be named according to the following conventions.
            The preferred form for variable names is all lower case
            letters and words separated with dots
            (variable.name), but variableName
            is also accepted;
            function names have initial capital letters and no dots
            (FunctionName);
            constants are named like functions but with an initial
            k.
            #https://google.github.io/styleguide/Rguide.xml#assignment
          
    


11. What is the value of x * y?
```{r}
y <- c(1, 2)
x <- y
y <- 3
x * y      # [3, 6]
```

## Factors