---
title: "Programmieren mit R: Seminararbeit_1"
author:
- Daniyar Akhmetov 
- Marcelo Rainho Avila
- Xuan Son Le
date: "Abgabedatum: 14/11/2017"
fontsize: 12
lof: false
graphics: true
documentclass: article
output: 
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part I (3 Points)  

```{r, include=FALSE, echo=FALSE}
rm(list = ls())     # remove all previous variables
```

1. *What are the atomic vector types in R?*
   *Explain which value they can take and give an example!*
2. *What is the difference between generic and atomic vectors?*
3. *Explain the following statement:* 
    *A data frame is a list, but not every list is a data frame.*

**Answer:**  
  
1. Atomic vectors are linear vectors (one-dimensional), which consists of values 
of the same type.

**Atomic vectors** | **Values**                          | **Examples**
-------------------|-------------------------------------|---------------
logical            | TRUE/FALSE                          | c(TRUE,FALSE)
integer            | integer numbers                     | c(1L, 10L, 100L)
numeric            | real numbers                        | c(-26, 1.3, -0.25)
complex            | complex numbers a+bi                | c(-2+3i, -4i, 0i)
character          | letters or words                    | c("Hello","R","How are you")
raw                | raw bytes (as pairs of hex digits)  | as.raw(255) = ff

2. Generic vectors (lists) are also linear vectors (one-dimensional), 
but can contain objects of different types. Example of a list: 
```{r}
x <- list(1:10,"a", c(TRUE,FALSE), c(1 + 2i,3i))
```

3. A data frame is a list of generic vectors of the same length and therefore has a two dimensional structure based on rows and columns. Example of a data frame:
```{r}
df <- data.frame(Name = c("M","X","D"), 
                 Country = c('B','V','R'), 
                 hasGlasses = c(FALSE,TRUE,TRUE))
df
```
Data frame is a list of lists with the same length. Each list from a data frame 
represents a column or a row of a data frame. But a list can not be a data frame 
because lists are linear vectors in one dimensional space. 
That's why we have the statement.

#Part II (7 Points) 
*Explain each line of the following code. In addition, discuss what the output of identical(a, b) will be. Check the help files for the functions set.seed, identical, rnorm and cumsum.*
```{r}
  # ensure that results from a random generator are reproducible
set.seed(1)
  # generate a vector of normally distributed random numbers
largeVector <- rnorm(1e7, mean = 5, sd = 10) 
  # define a and b
a <- cumsum(largeVector)[1:100] 
b <- cumsum(largeVector[1:100])
  # is a equal to b?
identical(a, b) 
```
**Answer:**  
  
* Because we ran the function set.seed before, the result of randomly choosen largeVector is reproducible for a and b. This means we are now working with the only one largeVector. 
* A *cumulative sum (cumsum)* is a sequence of partial sums of a given sequence.  
For instance:   
cumsum(largeVector)  
= cumsum(c(k1,k2,k3,...,kn))  
= k1  k1+k2  k1+k2+k3  ...  k1+k2+k3+..+kn  
  
* In a we firstly have a cumulative sum of largeVector and then take the first 100 elements from it:  
a = cumsum(largeVector)[1:100]  
a = (k1  k1+k2  k1+k2+k3 ...  k1+k2+k3+...+kn)[1:100]  
a =  k1  k1+k2  k1+k2+k3 ...  k1+k2+k3+...+k100  
  
* In b we firstly take the first 100 elements from largeVector and then calculate the cumulative sum of these 100 elements.  
b = cumsum(largeVector[1:100])  
b = cumsum(k1 k2 k3 ... k100)  
b = k1  k1+k2  k1+k2+k3 ...  k1+k2+k3+...+k100  

*system.time returns the time elapsed for the computation. Explain the differences:*  

**Answer:**    
```{r}
system.time(cumsum(largeVector)[1:100])
```
  
```{r}
system.time(cumsum(largeVector[1:100]))
```

\newpage

#Part III (20 Points)


*Conduct a regression analysis and report your findings. The report should contain the following sections:*

* *data import* 
* *descriptive statistics and data validation*
* *identification of relevant regressors*
* *fitting a regression model*
* *discussion of model fit, e.g. goodness of fit, significance of regressors*
* *interpreting the model*

**Answer:**  
**Logistic_Regression_Titanic:**  

## Data import 

Firstly we need to import the necessary libraries and dataset. 
```{r loading, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra) # install.packages("gridExtra") if needed

load(file = "./datasets/titanic.Rdata")
```
  
## Descriptive statistics and data validation
Let's take a look at the data to get a general overview of it
```{r}
  # See which variables are given in the data set through a small sample of the dataset
head(titanic,5)
```

Let's summarize the  variables 
```{r}
str(titanic)
```

**variable**| **description**                    | **Notes**
------------|------------------------------------|-----------------------
pclass      | Passenger Class                    | factor with 3 levels
survived    | Survival                           | labelled vector
name        | Passenger Name                     | labelled vector
sex         | Passenger Sex                      | factor with 2 levels
age         | Passenger Age                      | labelled vector
sibsp       | Number of Siblings/Spouses Aboard  | labelled vector
parch       | Number of Parents/Children Aboard  | labelled vector
ticket      | Ticket Number                      | labelled vector
fare        | Passenger Fare                     | labelled vector
cabin       | Cabin                              | factor with 187 levels
embarked    | Port of Embarkation                | factor with 3 levels
boat        | Lifeboat                           | factor with 28 levels
body        | Body Identification Number         | labelled vector
home.dest   | Home/Destination                   | labelled vector


Let's output a descriptive statistic of all variables for an overview of the data
```{r}
summary(titanic)
```



Since data frame aims to describes the survival status of individual passengers on the Titanic dependent on many different features. Let's take a look at some useful graphics. 

```{r, fig.width=3,fig.height=1.5}
  # Transform some variables, which are given as a vector
titanic$survived <- factor(titanic$survived)
titanic$age <- as.numeric(titanic$age)
titanic$fare <- as.numeric(titanic$fare)
titanic$sibsp <- as.numeric(titanic$sibsp)
titanic$parch <- as.numeric(titanic$parch)


ggplot(titanic, aes(survived, fill = survived)) + geom_bar() + 
    scale_fill_discrete(breaks = c("0", "1"),labels = c("No", "Yes")) + 
    labs(subtitle = "Frequency of survival")
```

```{r fig.height=6, fig.width=8}
  # Let's see how survival depends on other variables
p1 <- ggplot(titanic, aes(x = sex,fill = survived)) + 
    geom_bar() + scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) + 
    labs(subtitle = "Survival by Sex", y = "No. of passengers", x = " ")

p2 <- ggplot(titanic, aes(pclass,fill = survived)) + 
    geom_bar() + scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) + 
    labs(subtitle = "Survival by Passenger class", y = "No. of passengers", x = " ") + 
    ylim(0,800)
    
p3 <- ggplot(titanic, aes(sibsp,fill = survived)) + 
    geom_bar() + scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) + 
    labs(subtitle = "Survival by Number of\nSiblings/Spouses Aboard ", 
         y = "No. of passengers", x = " ")

p4 <- ggplot(titanic, aes(parch,fill = survived)) + 
    geom_bar() + scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) + 
    labs(subtitle = "Survival by Number of\nParents/Children Aboard ", 
         y = "No. of passengers", x = " ")

p5 <- ggplot(titanic, aes(age,fill = survived)) + geom_histogram(bins = 10) +
    scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) +
    labs(subtitle = "Survival by Age", y = "No. of passengers", x = " ")

p6 <- ggplot(titanic, aes(fare,fill = survived)) + geom_histogram(bins = 10) +
    scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes")) +
    labs(subtitle = "Survival by Age", y = "No. of passengers", x = " ")

grid.arrange(p1,p2,p3,p4,p5,p6,ncol = 2)

```

Let's create some more interesting graphs  
```{r}
  # 
ggplot(titanic, aes(x = interaction(sex,pclass), fill = survived)) + geom_bar() + 
    scale_fill_discrete(breaks = c("0", "1"), labels = c("No", "Yes"))
```

So after getting the first overlook of our data set as well as the given variables, let's take a look at possible missing values in the data set
```{r}
any(is.na(titanic))
sapply(titanic, function(x) sum(is.na(x)))
```

Some first annotations:  
* The variable *Age* contains about 20% missing values. We assume that Age does have influence on predicting the survival, so we will try to handle these missing values later on.  
* The varibale *body* contains about 91% missing values. It is quite a high amount. Probably we would drop this variable out of our model.  
* The amounts of missing vales of *fare* and *embarked* are very small. Maybe we would just drop the concerned passengers from the data set or replace the missing values by one certain value.  



##3. Identification of relevant regressors


Let's clean our data due to unnecessary information: 
Since some variables do not to contain a lot useful information to our model, we would drop them from our model. These are: *nam, cabin, ticket, boat, body, home.dest*
```{r}
titanic = titanic[ , !(names(titanic) %in% c("name","cabin","ticket","boat","body","home.dest"))]
```
  
Missing values also need to be handled  
* Variable *age*: Replace missing values by the mean *age*  
```{r}
titanic$age[is.na(titanic$age)] <- mean(titanic$age,na.rm = TRUE)
```
     
* Variable *fare*: Replace 1 missing value by the mean *fare* in the concerned *pclass*.  
```{r}
titanic$fare[is.na(titanic$fare)] <- mean(titanic$fare,na.rm = TRUE)
```


* Variable *embarked*: Replace 2 missing values by the most frequent *embarked* value.  
```{r}
summary(titanic$embarked)
titanic$embarked[is.na(titanic$embarked)] <- "Southampton"
```

Some variables also need to be relabeled, so that the values (for example 1,2,3) should not be treated as number but as category.  
```{r}
titanic$pclass <- factor(titanic$pclass, 
                         levels = c("1st","2nd","3rd"), 
                         labels = c(1,2,3))
titanic$sex <- factor(titanic$sex, 
                      levels = c("male","female"), 
                      labels = c(0,1))
titanic$embarked <- factor(titanic$embarked, 
                           levels = c("Cherbourg", "Queenstown", "Southampton"), 
                           labels = c(1, 2, 3)) 
```
  
Before we start modelling, let's check once again, whether all missing values are handled.  
```{r}
any(is.na(titanic))
```
  
**4. Fitting a regression model**  
Let's create a train and test data set
```{r}
titanic <- titanic[sample(1:nrow(titanic)),] # shuffle the rows
train <- titanic[1:1000,] # get the first 1000 passengers for training data
test <- titanic[1001:1309,] # the rest passengers for test data
```

Let's fit the logistic regression model.  
```{r}
model_full <- glm(survived ~. , family = binomial, data = train) # model with all filtered features
model_opt <- step(model_full) 
summary(model_opt)
```

**5. Discussion of model fit, e.g. goodness of fit, significance of regressors**

```{r}
  # Predicting the test results
test_predict <- predict(model_opt,type = 'response',
                        newdata = test[-2]) # apply the model to the test data
survived_predict <- ifelse(test_predict > 0.5,1,0) 
confusion_matrix <- table(test[,2],survived_predict) # create confusion matrix
confusion_matrix

accuracy <- mean(survived_predict == test$survived)*100
print(paste("Model accuracy is ", accuracy, "%"))


```


**6. Interpreting the model**





Alternative?:

load(file = "./datasets/miete.Rdata")

miete$bez <- as.integer(miete$bez)
miete$wohngut <- factor(miete$wohngut)
miete$wohnbest <- factor(miete$wohnbest)
miete$ww0 <- factor(miete$ww0)
miete$zh0 <- factor(miete$zh0)
miete$badkach0 <- factor(miete$badkach0)
miete$badextra <- factor(miete$badextra)
miete$kueche <- factor(miete$kueche)

Function defines normal/good/best living location
myfunction <- function(x,y){
 if (x == '0' && y == '0') {
     return(1)
 } else if (x == 1) {
     return(2)
 } else if (y == 1) {
     return(3)
 }
}

Apply defined function to wohngut and wohnbest
wohnlage <- mapply(myfunction, miete$wohngut, miete$wohnbest)

Replace variables wohngut and wohnbest by new varibale called wohnlage
miete <- data.frame(miete[1:6], wohnlage, miete[9:13])

Factorise variable wohnlage: 1 for normal, 2 for good and 3 for best
miete$wohnlage <- factor(miete$wohnlage)
levels(miete$wohnlage) <- list(normal = 1, gut = 2, best = 3)
str(miete)

max_model <- lm(formula = nm ~ nmqm + wfl + rooms + bj + bez + wohnlage + ww0 + zh0 + badkach0 +
              badextra + kueche, data = miete)
best_model <- step(max_model, direction = "backward")
max_model
best_model








